Marginalization of photometric redshift uncertainties

### 2020-07-21 Paper/paper.tex
Updates from Overleaf

\newcommand{\as}[1]{{\textcolor{blue}{[AS: #1]}}}
\newcommand{\an}[1]{{\textcolor{magenta}{[AN: #1]}}}
\newcommand{\da}[1]{{\textcolor{red}{[DA: #1]}}}

@@

\abstract{We present a new method for marginalization over uncertainties in redshift distributions of tomographic bins applicable to current and coming photometric galaxy surveys. We allow for arbitrary deviations from the best guess $N(z)$ governed by a general covariance matrix describing incompleteness in our knowledge of redshift distributions. In principle, this is marginalization over potentially hundreds of new parameters describing potential deviations as a function of redshift and tomographic bin. However, by Taylor expanding theory predictions around a fiducial model, we can perform the marginalization analytically, resulting in a modified data covariance matrix. We demonstrate  this method by applying to the galaxy clustering measurements from the Year One of Hyper SupremeCam data demonstrate how we can marginalize over sample-variance of the PZ calibration sample and a large general systematic uncertainty in photometric determination methods. We find that this method gives results that are comparable to the ad-hoc shift and width parameterizations and show that the only deviations that matter are smooth deviations. 
}

@@

  \section{Introduction}\label{sec:intro}\label{sec:intro}
    Photometric galaxy surveys  have the potential of  being transformative in our understanding of the Universe by measuring millions and soon billions of galaxies on the sky. These catalogs can be used to constrain cosmological parameters through measurements of clustering and weak lensing as demonstrated by numerous surveys that have completed observations including  Sloan Digital Sky Survey \cite{astro-ph/0006396}, PanSTARRS \cite{2010SPIE.7733E..0EK}, Kilo-degree Survey \cite{1206.1254}, Dark Energy Survey \cite{1601.00329}, Hyper Suprime-Cam survey \cite{2012SPIE.8446E..0ZM}, and will play a crucial role in the upcoming Vera Rubin Observatory LSST Survey \cite{0912.0201}.

    It has been long recognized that the accuracy of photometric surveys relies on the accuracy of photometric redshift distributions and that these are likely to continue to be a dominant source of systematic uncertainty. The crucial quantity is the redshift distribution $N(z)$, which is simply the mean number density of galaxies as a function of redshift for each tomographic sample. It is known that the shape of $N(z)$ delivered by the current photometric redshift codes and subsequent calibrations is insufficient for retiring this source of systematic error completely \cite{1809.01669}. Improving photometric redshift techniques and calibrations is an area of active research \cite{2006A&A...457..841I,2008ApJ...684...88N,2012MNRAS.423..909C,2013MNRAS.431.1547B,2016MNRAS.460.4258L,2019ApJ...877..117H,2018MNRAS.478..592H,2019MNRAS.489..820B,2020MNRAS.491.4768R,2019ApJ...881...80L,2019MNRAS.483.2487J,2019MNRAS.483.2801S,2019arXiv191007127A,2020A&A...637A.100W,2004.09542}, but for the foreseable future we will need to rely on marginalization to control residual systematic errors.

    The traditional approach is to add nuissiance parameters that shift the best guess $N(z)$ or additionally change its width or a relative contribution from a secondary peak in $N(z)$. This usually works well, but in additional to adding potentially large number of nuissance parameters suffers from model completeness problem: how do we know that any given parameterization encompasses all possible ways in which our best-guess $N(z)$ can be wrong?

    In this paper we propose a new technique that performs a marginalization around all possible functional deviations from the best fit $N(z)$. The space of possible deviations in $N(z)$ is described as a Gaussian function and under some controlled approximations, the integrals can be performed analytically. This results in a simple change to the data covariance matrix with a simple explanation: directions in the space of predictions that are degenerate with changes in $N(z)$ are given large variance, which means that information in these direction is not used for inferring the parameters of interest. This then marginalizes over $N(z)$ uncertainty in a robust manner that does not add new niussance parameters.

    We apply this method to the HSC data as an example and  discuss some of the advantages, limitations and possible extensions of this model.

  \section{Theory}\label{sec:theory}
    \subsection{Redshift distribution uncertainties}\label{ssec:theory.nz}
      Most cosmological analyses involve performing Bayesian parameter inference on a posterior distribution $p(\vec{\theta} |\vd) \propto p(\vd|\vec{\theta}) p(\vec{\theta})$, where $\vd$ is a vector of data points and $\vec{\theta}$ is a set of parameters descring the underlying model.

      In the case of tomographic analyses of galaxy clustering and large-scale structure, $\vd$ traditionally contains correlation function or power spectrum measurements between different tracers (e.g. galaxy ellipticities or overdensity) and different redshift bins\footnote{We use the term ``redshift bin'' here to denote any given galaxy sample, whether or not it has actually been selected by binning those galaxies into intervals of some redshift estimate.}, and $\vec{\theta}$ includes both cosmological and nuisance parameters needed to produce a forward model of $\vd$ that we will call $\vt(\vec{\theta})$. Due to the central limit theorem it is often accurate enough to assume that the likelihood is Gaussian, taking the form
      \begin{equation}
        \mathcal{L} \equiv p({\bf d}|\vec{\theta}) = \frac{\exp\left [-\frac{1}{2}(\vd-\vt(\vec{\theta}))^T{\sf C}^{-1}(\vd-\vt(\vec{\theta})) \right]}{\sqrt{\det{2\pi{\sf C}}}}, \label{eq:like}
     \end{equation}
      where ${\sf C}$ is the covariance matrix of $\vd$. The information contained in the parameter dependence of ${\sf C}$ can be neglected \citep{1811.11584}, and therefore the normalization factor $\sqrt{2\pi\det{\sf C}}$ can be ignored.

      We can make significant progress without specifying explicitly how the theory is calculated given the parameters. It suffices to say that, given a set of cosmological parameters, which specify the evolution of the universe and the matter inhomogeneities as a function of scale and redshift, a set of astrophysical parameters (e.g. a particular bias or intrinsic alignment model), and the redshift distributions $N(z)$ of the different redshift bins, we can integrate these into theory predictions $\vt=\vt(\vec{\theta})$. Our main concern in this paper is the way in which the significant uncertainties in $N(z)$, which can potentially dominate the total systematic error budget, should be parametrized and marginalized over.

      The traditional approach (e.g. \cite{2018PhRvD..98d3526A,1912.08209}) has been to parametrize departures from a given ``educated guess'' of the redshift distribution that are likely to describe the main modes in which the underlying uncertainty propagates into the theory vector. This has usually been done by introducing ``shift'' ($\Delta z$) and/or ``width'' ($z_w$) nuisance parameters, in terms of which the fiducial redshift distribution for the $i$-th redshift bin $\bar{N}_i(z)$ is modified as
      \begin{equation}
        N_{i}(z) \propto \bar{N}_{i}\left(z_{c,i} + (1 + z_{w, i})(z-z_{c, i}) + \Delta z_{i}\right), \label{eq:photo-z-model}
      \end{equation}
      where $z_{c,i}$ is the mean of the fiducial distribution and acts as a sensible pivot to define width changes. $\Delta z_i$ and $z_{w,i}$ for each bin are then inferred and marginalized over together with all other parameters, with priors on them based on calibration uncertainties.

      This approach explicitly marginalizes over two effects that are potentially the two most important systematics stemming from redshift distribution uncertainty: the width of the lens redshift distribution affects the amplitude and shape of the clustering power spectrum, and a shift in the distribution of sources influences the support of the lensing kernel and the amplitude of the shear two-point function. However, this methods suffers from two shortcomings. First, it is fundamentally ad-hoc: this particular parametric form, while intuitively the most important, is not directly mapped onto the main modes of uncertainty in photometric redshift estimators, and there is no rigorous proof that the parametrization is complete in the sense that it captures all relevant effects. The second problem is that it is computationally expensive. By adding one or two additional nuisance parameters per redshift bin, the dimensionality of the model parameter space increases significantly. For instance, in the analysis of \cite{1912.08209}, redshift uncertainty parameters accounted for approximately half of all model parameters. The methodology described in the next section aims to address both of these shortcomings.

    \subsection{A new approach}\label{ssec:theory.nz_new}
      Let us consider a discretized description of the $N_i(z)$ as a sum over basis functions $\phi_\alpha(z)$:
      \begin{equation}
        N_i(z)=\sum_i N^\alpha_i\phi_\alpha(z).
      \end{equation}
      For simplicity here we will just treat $N_i(z)$ as a histogram, and therefore in our case the $\phi_\alpha$ are just top-hat functions centered on an equi-spaced grid of redshifts $z_\alpha$. The formalism described below is however valid for any choice of basis functions. In particular, it is likely that future photometric estimators will provide their uncertainty in terms of a several important components and a corresponding covariance matrix.    We describe the full set of redshift distributions by the vector $\vN\equiv\{N^\alpha_i\,\,\,\forall i,\alpha\}$. In what follows, we will assume that $\vN$ has units of galaxy ``counts'' \as{Do we actually need to say this?}. The actual normalization of the $N(z)$ is irrelevant, since it only enters the theory prediction for the power spectrum as a normalized redshift probability distribution (see Eq. \ref{eq:cell_gg_limber}).

@@

      Our data is therefore made up of the combination ${\bf d}=(\hat{\bf c},\hat{\bf N})$, where $\hat{\bf c}$ is a set of two-point function measurements. The model parameters are $\vec{\theta}=({\bf q},\vN)$, where $\vN$ are the redshift distribution coefficients, and ${\bf q}$ contains all other cosmological, astrophysical and nuisance parameters. The posterior distribution is therefore given by:
      \begin{align}
        p(\vec{\theta}|{\bf d})
        &=p({\bf q},\vN|\hat{\bf c},\hat{\bf N})\\\label{eq:posterior_full}
        &=p(\hat{\bf c}|{\bf q},\vN)p(\hat{\bf N}|\vN)p(\vN)p({\bf q})\\
        &=p(q)\frac{\exp\left[-\frac{1}{2}(\chi^2_c+\chi^2_N+\chi^2_P)\right]}{\sqrt{{\rm det}(2\pi{\sf C}_c){\rm det}(2\pi{\sf C}_N){\rm det}(2\pi{\sf C}_P)}},
      \end{align}
      where
      \begin{align}
        \chi^2_c&=\left(\hat{\bf c}-{\bf t}\right)^T{\sf C}_c^{-1}\left(\hat{\bf c}-{\bf t}\right)\\
        \chi^2_N&=\left(\hat{\bf N}-\vN\right)^T{\sf C}_N^{-1}\left(\hat{\bf N}-\vN\right)\\
        \chi^2_P&=\left(\vN-\vN_P\right)^T{\sf C}_P^{-1}\left(\vN-\vN_P\right),
      \end{align}
      and ${\bf t}(q,\vN)$ is the theoretical prediction for $\hat{\bf c}$.

      In Eq. \ref{eq:posterior_full} we have assumed that $\hat{\bf c}$ and $\hat{\vN}$ are independent at the likelihood level, and that the likelihood of $\hat{\vN}$ is independent of ${\bf q}$. The validity of these assumptions should be studied in detail, especially in cases where $\vN$ is calibrated using a spectroscopic sample with significant spatial overlap with the data under study (through a clustering redshifts approach \cite{2008ApJ...684...88N} or otherwise). The method outlined here is still applicable if these assumptions are dropped, albeit with a different expression for the modified covariance in Eq. \ref{eq:main2}. We have also assumed that the deviations from the best guess redshift distribution $\hat{\vN}$ are Gaussianly distributed. While this is likely not true in detail for the current $N(z)$, it will likely become a better approximation in the future as our data become more precise. Moreover, it is also unlikely to  matter much in practice, as long as $C_N$ captures the relevant directions of uncertainty.

      To proceed, let us start by considering the combination $\chi^2_N+\chi^2_P$. After completing squares for $\vN$, this can be written as:
      \begin{equation}
        \chi^2_N+\chi^2_P\equiv\chi^2_{\bar{N}}=\left(\vN-\tilde{\vN}\right)^T{\sf P}^{-1}\left(\vN-\tilde{\vN}\right)+K_N,
      \end{equation}
      where we have defined the \emph{smoothed mean} $\bar{\vN}$ and combined prior covariance ${\sf P}$ as:
      \begin{align}\label{eq:priorcov}
        \bar{\vN}\equiv{\sf P}({\sf C}_N^{-1}\hat{\bf N}+{\sf C}_P^{-1}\vN_P),\hspace{12pt} {\sf P}^{-1}\equiv{\sf C}_N^{-1}+{\sf C}_P^{-1},
      \end{align}
      and $K_N$ is independent of the model parameters, and given by
      \begin{equation}
        K_N\equiv\hat{\bf N}^T{\sf C}^{-1}_N\hat{\bf N}+\vN_P^T{\sf C}_P^{-1}\vN_P-\bar{\vN}^T{\sf P}^{-1}\bar{\vN}.
      \end{equation}
      The full posterior thus takes the form:
      \begin{align}
        p({\bf q},\vN|\hat{\bf c},\hat{\bf N})&=Q\,p({\bf q})\exp\left[-\frac{1}{2}(\hat{\bf c}-{\bf t})^T{\sf C}^{-1}_c(\hat{\bf c}-{\bf t})-\frac{1}{2}(\vN-\bar{\vN})^T{\sf P}^{-1}(\vN-\bar{\vN})\right],\\
        Q&\equiv\frac{e^{-K_N/2}}{\sqrt{{\rm det}(2\pi{\sf C}_c){\rm det}(2\pi{\sf C}_N){\rm det}(2\pi{\sf C}_P)}},
      \end{align}

%We will describe our uncertainty in $\vN$ as a Gaussian
%\begin{equation}
%  p(\vN) \propto \exp\left[  (\vN - \bar{\vN})^T P^{-1} (\vN -\bar{\vN}) \right],
%\end{equation}
%where $\hat{\vN}$ is our best guess of redshift distribution $P$ is a matrix of allowed deviations. We will discuss later how $P$ can be determined, but at this point we note that it is a general positive-definite matrix, i.e not necessarily diagonal). Other possible parameterizations are sub-summed in this approach. For example, a general shift-width is allowed with the probability $p(\vN(\Delta z_i, z_{w,i}))$.

      We can now write a likelihood for $\hat{\bf c}$ in which we explicitly marginalize out all the degrees of freedom associated with $\vN$:
      \begin{equation}
        \mathcal{L} \propto  \int \exp\left [-\frac{1}{2} ({\bf c}-\vt)^T C^{-1} ({\bf c}-\vt) -\frac{1}{2} (\vN - \bar{\vN})^T P^{-1} (\vN -\bar{\vN}) \right]  d^N\vN, \label{eq:like2}
      \end{equation}
      Note that in the equation above, $\vt=\vt({\bf q},\vN)$. It is the non-linear dependence of $\vt$ on $\vN$ that prevents us from performing what looks like a trivial integral analytically. In principle, we could solve this problem by making every single element of vector $\vN$ a part of an MCMC chain for ${\bf q}$ and additional hundreds of $\vN$ parameters. This might actually be possible in the future using methods that can sample very large parameter spaces, such as Hamiltonian Monte Carlo \cite{1987PhLB..195..216D}. Instead, faced with queue waiting times at NERSC we Taylor expand theory around $\bar{\vN}$:
      \begin{equation}
        \vt({\bf q},\vN) \simeq \vt({\bf q},\bar{\vN}) + {\sf T} \left(\vN - \bar{\vN} \right), \label{eq:taylor}
      \end{equation}
      where the matrix ${\sf T}$ is the gradient of ${\bf t}$ with respect to $\vN$
      \begin{equation}
        {\sf T}\equiv \left.\frac{d{\bf t}}{d\vN}\right|_{{\bf q},\bar{\vN}}
        \label{eq:Tmat}
      \end{equation}
      In other words it contains a full information about how all the correlation functions or power spectra respond to a small change in $N(z)$.

      Substituting Eq. \ref{eq:taylor} into Eq. \ref{eq:like2} results in an integral that is now quadratic in $\vN$ and can be performed analytically. After completing squares, and applying the Woodbury matrix identity, we find:
      \begin{equation}
        \mathcal{L} \propto \left[{\rm det}\left({\sf T}^T {\sf C}_c^{-1} {\sf T} +{\sf P}^{-1}\right)\right]^{-1/2} \exp\left [-\frac{1}{2} ({\bf c}-\vt)^T {\sf C}_M^{-1} ({\bf c}-\vt) \right], \label{eq:main1}
      \end{equation}
      where the  marginalized covariance is
      \begin{equation}
        {\sf C}_M = {\sf C}_c + {\sf T}{\sf P}{\sf T}^T \label{eq:main2}
      \end{equation}
      This  fascinatingly simple equation and the main result of this paper. This calculation should be understood as follows: for each deviation around $\hat{\vN}$, there is a corresponding deviation around $\vt$. These are directions in space of theory predictions that are perfectly degenerate with the changes in shape of $N(z)$.  Equation \ref{eq:main2}  increases the variance for those linear combinations commensurate with how far ${\sf P}$ allows them to go. The limit ${\sf P}\rightarrow\infty\mathbb{1}$ would correspond to completely projecting out those ``$N(z)$-sensitive'' modes from the data altogether.

>>>>

\newcommand{\as}[1]{{\textcolor{blue}{[AS: #1]}}}
\newcommand{\an}[1]{{\textcolor{magenta}{[AN: #1]}}}
\newcommand{\da}[1]{{\textcolor{red}{[DA: #1]}}}
\newcommand{\bh}[1]{{\textcolor{teal}{[BH: #1]}}}

@@

\abstract{We present a new method to marginalize over uncertainties in redshift distributions, $N(z)$, within tomographic cosmological analyses applicable to current and upcoming photometric galaxy surveys. We allow for arbitrary deviations from the best guess $N(z)$ governed by a general covariance matrix describing the uncertainty in our knowledge of redshift distributions. In principle, this is marginalization over potentially hundreds or thousands of new parameters describing potential deviations as a function of redshift and tomographic bin. However, by linearly expanding the theory predictions around a fiducial model, this marginalization can be performed analytically, resulting in a modified data covariance matrix that effectively downweights the modes of the data vector that are more sensitive to redshift distribution variations. We showcase this method by applying it to the galaxy clustering measurements from the Hyper Suprime-Cam first data release. We demonstrate how we can marginalize over sample-variance of the PZ calibration sample and a large general systematic uncertainty in photometric determination methods, and explore the impact of priors imposing smoothness in the redshift distributions. We find that this method gives results that are comparable to the ad-hoc shift and width parameterizations, and show that the only deviations that matter are smooth deviations \da{are we really doing this?}. 
}

@@

  \section{Introduction}\label{sec:intro}\label{sec:intro}
    Photometric galaxy surveys  have the potential to transform our understanding of the Universe by measuring the properties of millions and soon billions of galaxies on the sky. These catalogs can be used to constrain cosmological parameters from measurements of galaxy clustering and weak gravitational lensing as demonstrated by numerous surveys, including the Sloan Digital Sky Survey \cite{astro-ph/0006396}, PanSTARRS \cite{2010SPIE.7733E..0EK}, Kilo-degree Survey \cite{1206.1254}, Dark Energy Survey \cite{1601.00329}, Hyper Suprime-Cam survey (HSC) \cite{2012SPIE.8446E..0ZM}, and will play a crucial role in the upcoming Vera Rubin Observatory LSST Survey \cite{0912.0201} \da{We could also cite the actual analysis papers}.

    It has been long recognized that the constraining power of photometric surveys relies heavily on the accuracy of photometric redshift distributions and that this is likely to continue to be a dominant source of systematic uncertainty. The crucial quantity is the redshift distribution $N(z)$, which is simply the mean number of galaxies as a function of redshift for each tomographic sample. It is known that the shape of $N(z)$ delivered by the current photometric redshift codes and subsequent calibrations is insufficient for retiring this source of systematic error completely \cite{1809.01669}. Improving photometric redshift techniques and calibration methods for redshift distributions is an area of active research \cite{2006A&A...457..841I,2008ApJ...684...88N,2012MNRAS.423..909C,2013MNRAS.431.1547B,2016MNRAS.460.4258L,2019ApJ...877..117H,2018MNRAS.478..592H,2019MNRAS.489..820B,2020MNRAS.491.4768R,2019ApJ...881...80L,2019MNRAS.483.2487J,2019MNRAS.483.2801S,2019arXiv191007127A,2020A&A...637A.100W,2004.09542}, but for the foreseeable future, reliable methods to marginalize over uncertainties in the $N(z)$ well be crucial to control residual systematic errors.

    The traditional approach has been to add nuisance parameters that shift the mean of the best guess $N(z)$ or, additionally, change its width or a relative contribution from a secondary peak in $N(z)$. This approach has so far been sufficient but, in addition to adding a potentially large number of nuisance parameters, the method suffers from a model completeness problem: how do we know that any particular parametrization encompasses all possible ways in which our best-guess $N(z)$ can be wrong?

    In this paper we propose a new technique that performs a marginalization around all possible functional deviations from the best-guess $N(z)$. The space of possible deviations in $N(z)$ is described as a Gaussian function and, under some controlled approximations, these deviations can be marginalized over analytically. This results in a simple change to the data covariance matrix with a simple explanation: directions in the space of predictions that are degenerate with changes in $N(z)$ are given large variance, which means that information in these direction is not used for inferring the parameters of interest. The method then marginalizes over the $N(z)$ uncertainty in a robust manner, without including a large number of new nuisance parameters.

    We apply this method to the HSC first public data release as an example, and discuss some of the advantages, limitations and possible extensions of this model.

    \da{The paper is structured as follows...}

  \section{Theory}\label{sec:theory}
    \subsection{Redshift distribution uncertainties}\label{ssec:theory.nz}
      Most cosmological analyses involve performing Bayesian parameter inference on a posterior distribution $p(\vec{\theta} |\vd) \propto p(\vd|\vec{\theta}) p(\vec{\theta})$, where $\vd$ is a vector of data points and $\vec{\theta}$ is a set of parameters describing the underlying model.

      In the case of tomographic large-scale structure analyses, $\vd$ traditionally contains correlation function or power spectrum measurements between different tracers (e.g. galaxy ellipticities or overdensity) and different redshift bins\footnote{We use the term ``redshift bin'' here to denote any given galaxy sample, whether or not it has actually been selected by binning those galaxies into intervals of some redshift estimate.}, and $\vec{\theta}$ includes both cosmological and nuisance parameters needed to produce a forward model of $\vd$ that we will call $\vt(\vec{\theta})$. Due to the central limit theorem it is often accurate enough to assume that the likelihood is Gaussian \cite{2018MNRAS.473.2355S}, taking the form
      \begin{equation}
        \mathcal{L} \equiv p({\bf d}|\vec{\theta}) = \frac{\exp\left [-\frac{1}{2}(\vd-\vt(\vec{\theta}))^T{\sf C}^{-1}(\vd-\vt(\vec{\theta})) \right]}{\sqrt{\det{2\pi{\sf C}}}}, \label{eq:like}
     \end{equation}
      where ${\sf C}$ is the covariance matrix of $\vd$. The information contained in the parameter dependence of ${\sf C}$ can be neglected \citep{1811.11584}, and therefore the normalization factor $\sqrt{2\pi\det{\sf C}}$ can be ignored.

      We can make significant progress without specifying explicitly how the theory is calculated given the parameters. It suffices to say that, given a set of cosmological parameters, which specify the expansion history of the universe and the growth of matter inhomogeneities as a function of scale and redshift, a set of astrophysical parameters (e.g. a particular bias or intrinsic alignment model), and the redshift distributions $N_i(z)$ of the different redshift bins, we can integrate these into theory predictions $\vt=\vt(\vec{\theta})$. Our main concern in this paper is the way in which the significant uncertainties in $N(z)$, which can potentially dominate the total systematic error budget, should be parametrized and marginalized over.

      The traditional approach (e.g. \cite{2018PhRvD..98d3526A,2019PASJ...71...43H,1912.08209,2020A&A...637A.100W}) has been to parametrize departures from a given ``educated guess'' of the redshift distribution that are likely to describe the main modes in which the underlying uncertainty propagates into the theory vector. This has usually been done by introducing ``shift'' ($\Delta z$) and/or ``width'' ($z_w$) nuisance parameters, in terms of which the fiducial redshift distribution for the $i$-th redshift bin $\bar{N}_i(z)$ is modified as
      \begin{equation}
        N_{i}(z) \propto \bar{N}_{i}\left(z_{c,i} + (1 + z_{w, i})(z-z_{c, i}) + \Delta z_{i}\right), \label{eq:photo-z-model}
      \end{equation}
      where $z_{c,i}$ is the mean of the fiducial distribution and acts as a sensible pivot to define width changes. $\Delta z_i$ and $z_{w,i}$ for each bin are then inferred and marginalized over together with all other parameters, with priors on them based on calibration uncertainties.

      This approach explicitly marginalizes over two effects that are potentially the two most important systematics stemming from redshift distribution uncertainty: the width of the lens redshift distribution affects the amplitude and shape of the clustering power spectrum, and a shift in the distribution of sources influences the support of the lensing kernel and the amplitude of the shear two-point function. However, this methods suffers from two shortcomings. First, it is fundamentally ad-hoc: while this particular parametric form, intuitively characterizes the most important effects, is not directly mapped onto the main modes of uncertainty in photometric redshift estimators, and there is no rigorous proof that the parametrization is complete in the sense that it captures all relevant effects. The second problem is that it is computationally expensive. By adding one or two additional nuisance parameters per redshift bin, the dimensionality of the model parameter space increases significantly. For instance, in the analysis of \cite{1912.08209}, redshift uncertainty parameters accounted for approximately half of all model parameters. The methodology described in the next section aims to address both of these shortcomings.

    \subsection{A new approach}\label{ssec:theory.nz_new}
      Let us consider a discretized description of the $N_i(z)$ as a sum over basis functions $\phi_\alpha(z)$:
      \begin{equation}
        N_i(z)=\sum_i N^\alpha_i\phi_\alpha(z).
      \end{equation}
      For simplicity here we will just treat $N_i(z)$ as a histogram, and therefore in our case the $\phi_\alpha$ are just top-hat functions centered on an equi-spaced grid of redshifts $z_\alpha$. The formalism described below is however valid for any choice of basis functions. In particular, it is likely that future photometric estimators will provide their uncertainty in terms of a several important components and a corresponding covariance matrix.    We describe the full set of redshift distributions by the vector $\vN\equiv\{N^\alpha_i\,\,\,\forall i,\alpha\}$. In what follows, we will assume that $\vN$ has units of galaxy ``counts'' \as{Do we actually need to say this?} \da{This only affects the specific form of the shot-noise covariance Eq. \ref{eq:cov_nz_shot}}. The actual normalization of the $N(z)$ is irrelevant, since it only enters the theory prediction for the power spectrum as a normalized redshift probability distribution (see Eq. \ref{eq:cell_gg_limber}).

@@

      Our data is therefore made up of the combination ${\bf d}=(\hat{\bf c},\hat{\bf N})$, where $\hat{\bf c}$ is a set of two-point function measurements. The model parameters are $\vec{\theta}=({\bf q},\vN)$, where $\vN$ are the redshift distribution coefficients, and ${\bf q}$ contains all other cosmological, astrophysical and nuisance parameters. The posterior distribution is therefore given by:
      \begin{align}
        p(\vec{\theta}|{\bf d})
        &=p({\bf q},\vN|\hat{\bf c},\hat{\bf N})\\\label{eq:posterior_full}
        &\propto p(\hat{\bf c}|{\bf q},\vN)p(\hat{\bf N}|\vN)p(\vN)p({\bf q})\\
        &=p(q)\frac{\exp\left[-\frac{1}{2}(\chi^2_c+\chi^2_N+\chi^2_P)\right]}{\sqrt{{\rm det}(2\pi{\sf C}_c){\rm det}(2\pi{\sf C}_N){\rm det}(2\pi{\sf C}_P)}},
      \end{align}
      where
      \begin{align}
        \chi^2_c&=\left(\hat{\bf c}-{\bf t}\right)^T{\sf C}_c^{-1}\left(\hat{\bf c}-{\bf t}\right)\\
        \chi^2_N&=\left(\hat{\bf N}-\vN\right)^T{\sf C}_N^{-1}\left(\hat{\bf N}-\vN\right)\\
        \chi^2_P&=\left(\vN-\vN_P\right)^T{\sf C}_P^{-1}\left(\vN-\vN_P\right),
      \end{align}
      and ${\bf t}(q,\vN)$ is the theoretical prediction for $\hat{\bf c}$.

      In Eq. \ref{eq:posterior_full} we have assumed that $\hat{\bf c}$ and $\hat{\vN}$ are independent at the likelihood level, and that the likelihood of $\hat{\vN}$ is independent of ${\bf q}$. The validity of these assumptions should be studied in detail, especially in cases where $\vN$ is calibrated using a spectroscopic sample with significant spatial overlap with the data under study (through a clustering redshifts approach \cite{2008ApJ...684...88N} or otherwise). The method outlined here is still applicable if these assumptions are dropped, albeit with a different expression for the modified covariance in Eq. \ref{eq:main2}. We have also assumed that the deviations from the best guess redshift distribution $\hat{\vN}$ are Gaussianly distributed \cite{2004.09542}. While this is not true in detail for the current $N(z)$, it will likely become a better approximation in the future as data become more precise. Moreover, it is also unlikely to  matter much in practice, as long as $C_N$ captures the relevant directions of uncertainty.

      To proceed further, let us start by considering the combination $\chi^2_N+\chi^2_P$. After completing squares for $\vN$, this can be written as:
      \begin{equation}
        \chi^2_N+\chi^2_P\equiv\chi^2_{\bar{N}}=\left(\vN-\tilde{\vN}\right)^T{\sf P}^{-1}\left(\vN-\tilde{\vN}\right)+K_N,
      \end{equation}
      where we have defined the \emph{smoothed mean} $\bar{\vN}$ and combined prior covariance ${\sf P}$ as:
      \begin{align}\label{eq:priorcov}
        \bar{\vN}\equiv{\sf P}({\sf C}_N^{-1}\hat{\bf N}+{\sf C}_P^{-1}\vN_P),\hspace{12pt} {\sf P}^{-1}\equiv{\sf C}_N^{-1}+{\sf C}_P^{-1},
      \end{align}
      and $K_N$ is independent of the model parameters, and given by
      \begin{equation}
        K_N\equiv\hat{\bf N}^T{\sf C}^{-1}_N\hat{\bf N}+\vN_P^T{\sf C}_P^{-1}\vN_P-\bar{\vN}^T{\sf P}^{-1}\bar{\vN}.
      \end{equation}
      The full posterior thus takes the form:
      \begin{align}
        p({\bf q},\vN|\hat{\bf c},\hat{\bf N})&=Q\,p({\bf q})\exp\left[-\frac{1}{2}(\hat{\bf c}-{\bf t})^T{\sf C}^{-1}_c(\hat{\bf c}-{\bf t})-\frac{1}{2}(\vN-\bar{\vN})^T{\sf P}^{-1}(\vN-\bar{\vN})\right],\\
        Q&\equiv\frac{e^{-K_N/2}}{\sqrt{{\rm det}(2\pi{\sf C}_c){\rm det}(2\pi{\sf C}_N){\rm det}(2\pi{\sf C}_P)}},
      \end{align}

      We can now write a likelihood for $\hat{\bf c}$ in which we explicitly marginalize out all the degrees of freedom associated with $\vN$:
      \begin{equation}
        \mathcal{L} \propto  \int \exp\left [-\frac{1}{2} ({\bf c}-\vt)^T C^{-1} ({\bf c}-\vt) -\frac{1}{2} (\vN - \bar{\vN})^T P^{-1} (\vN -\bar{\vN}) \right]  d^N\vN, \label{eq:like2}
      \end{equation}
      Note that in the equation above, $\vt=\vt({\bf q},\vN)$. It is the non-linear dependence of $\vt$ on $\vN$ that prevents us from performing what looks like a trivial integral analytically. In principle, we could solve this problem by making every single element of vector $\vN$ a part of an MCMC chain for ${\bf q}$ and additional hundreds of $\vN$ parameters. This should be feasible using methods that can sample very large parameter spaces, such as Hamiltonian Monte Carlo \cite{1987PhLB..195..216D}. Instead, faced with queue waiting times at NERSC \da{I'm fine leaving this XD, but the referee or journal will complain} we can Taylor expand the theory in $\vN$ to first order around $\bar{\vN}$:
      \begin{equation}
        \vt({\bf q},\vN) \simeq \vt({\bf q},\bar{\vN}) + {\sf T} \left(\vN - \bar{\vN} \right), \label{eq:taylor}
      \end{equation}
      where the matrix ${\sf T}$ is the gradient of ${\bf t}$ with respect to $\vN$
      \begin{equation}
        {\sf T}\equiv \left.\frac{d{\bf t}}{d\vN}\right|_{{\bf q},\bar{\vN}}
        \label{eq:Tmat}
      \end{equation}
      In other words ${\sf T}$ contains a full information about how all the correlation functions or power spectra respond to a small change in $N_i(z)$.

      Substituting Eq. \ref{eq:taylor} into Eq. \ref{eq:like2} results in an integral that is now quadratic in $\vN$ and can be performed analytically. After completing squares, and applying the Woodbury matrix identity, we find:
      \begin{equation}
        \mathcal{L} \propto \left[{\rm det}\left({\sf T}^T {\sf C}_c^{-1} {\sf T} +{\sf P}^{-1}\right)\right]^{-1/2} \exp\left [-\frac{1}{2} ({\bf c}-\vt)^T {\sf C}_M^{-1} ({\bf c}-\vt) \right], \label{eq:main1}
      \end{equation}
      where the  marginalized covariance is
      \begin{equation}
        {\sf C}_M = {\sf C}_c + {\sf T}{\sf P}{\sf T}^T \label{eq:main2}
      \end{equation}
      This is a fascinatingly simple equation and the main result of this paper. This calculation can be understood as follows: for each deviation around $\bar{\vN}$, there is a corresponding deviation around $\vt$. These are directions in space of theory predictions that are perfectly degenerate with the changes in shape of $N_i(z)$.  Equation \ref{eq:main2}  increases the variance for those linear combinations commensurate with how far ${\sf P}$ allows them to go. The limit ${\sf P}\rightarrow\infty\mathbb{1}$ would correspond to completely projecting out those ``$N(z)$-sensitive'' modes from the data altogether.

      Both the marginalized covariance ${\sf C}_M$, and the normalizing prefactor in Eq. \ref{eq:main1} depend in principle on ${\bf q}$ through the parameter dependence of ${\sf T}$. This implies that, in principle, ${\sf T}$ should be re-evaluated at every point in the MCMC chain. Since the calculation of ${\sf T}$ is expensive (by comparison with that of e.g. $\vt$) using standard methods, we will neglect this dependence and evaluate ${\sf T}$ at a fiducial set of parameters. We will however explore the impact of the choice of fiducial parameters on the final results. This should be a sufficiently good approximation for compact likelihoods, where parameter constraints are driven by the differences between $\hat{\bf c}$ and $\vt$. We will therefore pre-compute $T$ once, which then means that for a fixed data covariance matrix ${\sf C}_c$, the normalization term in equation \ref{eq:main1} is an irrelevant constant, and the marginalized covariance in the Eq. \ref{eq:main2} also needs to be evaluated only once.

      The procedure to obtain constraints from this marginalized likelihood would therefore proceed as follows:
      \begin{itemize}
        \item Solve the parameter inference problem once with fixed redshift distribution and a standard Gaussian likelihood for $\hat{\bf c}$ with covariance ${\sf C}_c$ to determine a sensible fiducial model.
        \item Calculate ${\sf T}$ and ${\sf C}_M$ at this fiducial model.
        \item Solve again the inference problem with a standard Gaussian likelihood using the modified covariance matrix ${\sf C}_M$ instead of ${\sf C}_c$. This will lead to broadened contours due to the marginalization over redshift distribution uncertainties.
      \end{itemize}
      If needed, one could repeat the inference at a refined fiducial model although in practice we found this to be unnecessary.
